Fermi and eROSITA bubbles

### 2021-12-09 paper.tex
update

\section{Methodology}
\subsection{Assumptions and Numerical Techniques}
\begin{enumerate}
  \item Numerical features:
    \begin{enumerate}
      \item GPU acceleration with \textsc{gamer-sr} code.
      \item Special relativistic hydrodynamics.
      \item We use a new algorithm \citep{tseng2021} to convert between\
            primitive and conserved variables, significantly reducing\
            numerical error caused by catastrophic cancellations that commonly occur\
            within the regions with high Mach number. i.e. jet-ISM interaction zones.
      \item To further alleviate the numerical error due to high Mach number,\
            \textsc{gamer-sr} also adaptively reduce the min-mod coefficient \citep{tseng2021}\
            inside the failed patch.\
            Doing so provides an elegant way to avoid the use of pressure/density floor,\
            which is unnatural but widely used in almost publicly available code.
    \end{enumerate}
  \item Assumptions on hydrodynamics:
    \begin{enumerate}
      \item Ideal fluid with TM EoS (\citealt{Taub}; \citealt{TM_EOS}).
    \end{enumerate}
  \item Assumptions on cosmic-ray:
    \begin{enumerate}
      \item We treat CRs as a second fluid and solve directly for the evolution of CR pressure $p_{\text{cr}}$\
            as a function of $\mathbf{r}$ and $t$.
      \item We did not model the CR energy spectrum.
      \item We neglected the cooling and heating processes of CRs, such as energy losses due to synchrotron\
            and inverse Compton emission, and reacceleration in shocks.
      \item We have assumed cosmic-ray is passive. (i.e. $p_{\text{cr}}\ll p_{\text{gas}}$)
      \item \textcolor{red}{We have ignored the $\mathbf{B}$ field within\
            the simulation box as the field inside\
            the bubbles should be weak due to adiabatic expansion, and thus the magnetic fields has\
            little effect on the overall dynamics.\
            We also assumed the magnetic field is highly entangled on small-scale (how small?), resulting in\
            the negligible cosmic-ray diffusion.}
    \end{enumerate}
  \item Assumptions on gravity:
    \begin{enumerate}
      \item We use the potential of isothermal slab, symmetric about $z=0$ and\
            supported by gas pressure and by under its own self-gravity,\
            to mimic the realistic potential of Galactic bulge.
      \item We still use Newtonian gravity, as the relativistic fluid ejected by the jet source\
            is quickly slowed down by heavy disk in a short time, and the relativistic fluid\
            accounts for a little minority of total mass inside the simulation box.
      \item The ISM disk and atmosphere are subjected to the fixed external potential\
            due to a disk bulge and dark matter halo.
      \item In addition to the gravitational interaction, we ignore other interactions between stars and gases.
      \item We also ignore the self-gravity of the ISM disk and of the atmosphere.
      \item We ignore the centrifugal force of Milky Way rotation acting on the bubbles.
      \item We use the potential of isothermal slab to mimic the gravitational potential due to the stellar bulge.
      \item The interface between cold ISM disk and atmosphere is parallel to the Galactic plane and is pressure balanced.
    \end{enumerate}
\end{enumerate}
We simulate 3D special relativistic hydrodynamics with passive CR injections from the GC using\
the special hydrodynamics GPU code \textsc{gamer-sr} \citep{tseng2021}.

@@

\end{enumerate}
\subsection{Jet injection}

\subsection{X-ray and Gamma-ray emission}
  \begin{enumerate}
    \item X-ray:
       \begin{enumerate}
         \item Thermal bremsstrahlung: The X-ray emissivity in an energy range 1.4–1.6 keV is calculated\
               using the MEKAL model (\citealt{Xray-1}; \citealt{Xray-2}; \citealt{Xray-3})\
               implemented in the utility XSPEC\citep{XSPEC}, assuming solar metallicity.
       \end{enumerate}
    \item Gamma-ray:
       \begin{enumerate}
         \item Leptonic process:\\
          The gamma-ray emission is produced by inverse Comptom scattering of the ISRF by CRe.
         \item Hadronic process:\\
          In the hadronic model, CRp undergo hadronic collisions with thermal gas protons\
          and produce $\gamma$-ray via pion decay. The volume emissivity of the emission can be written as
          \begin{equation}
             \epsilon \propto U_{\text{CRp}}n_{p}\sigma_{p}\kappa_{pp}.
          \end{equation}
       \end{enumerate}
  \end{enumerate}

@@

>>>>

\section{Methodology}
\subsection{Assumptions and Numerical Techniques}
  We used the GPU-accelerated special relativistic hydrodynamics AMR code (\textsc{gamer-sr}) developed at the\
  National Taiwan University\
  (Schive et al. \citeyear{gamer-1}, \citeyear{gamer-2}; \citeauthor{tseng2021} \citeyear{tseng2021}) to carry out\
  the simulations of Fermi and eROSITA bubbles. \textsc{gamer-sr} adopts a new algorithm \citep{tseng2021} to convert between\
  primitive and conserved variables, significantly reducing numerical error caused by catastrophic cancellations\
  that commonly occur within the regions with high Mach number. i.e. jet-ISM interaction zones.

  \textsc{gamer-sr} also adaptively and locally reduce the min-mod coefficient \citep{tseng2021} within the failed patch group.\
  Doing so provides an elegant way to avoid the use of pressure/density floor,\
  being unnatural but widely used in almost publicly available code.
  %to further alleviate the numerical error due to extremely high Mach number,\

  The simulation box is $14\times14\times28$ kpc, slightly larger than the size of eROISTA bubbles,\
  with a root grid $N=16\times16\times32$. We resolve the jet source located at the center of the simulation domain,\
  cells are refined to level 11, leading to a finest spatial resolution of 0.4 pc.\

  To save computational costs further, we restrict the refinement level is 7 within the cold disk so that\
  a molecular colud can be adequately resolved by approximately 30 cells along their diameter, 100 pc.

  We treat CRs as a second fluid and solve directly for the evolution of CR pressure $p_{\text{cr}}$\
  as a function of $\mathbf{r}$ and $t$.\
  We did not model the CR energy spectrum and neglect the cooling and heating processes of CRs,\
  such as energy losses due to synchrotron and inverse Compton emission, and reacceleration in shocks.

  We have assumed cosmic-ray is passive. (i.e. $p_{\text{cr}}\ll p_{\text{gas}}$)

  We have ignored the $\mathbf{B}$ field within\
  the simulation box as the field inside\
  the bubbles should be weak due to adiabatic expansion, and thus the magnetic fields has\
  little effect on the overall dynamics.\

  We also assumed the magnetic field is highly entangled on small-scale (how small?), resulting in\
  the negligible cosmic-ray diffusion.

  \textsc{gamer-sr} solves the mass and energy-momentum conservation laws of special relativistic ideal fluid\
  including CR advection, and dynamical coupling between the thermal gas and CRs. The governing equations\
  can be written as

\begin{enumerate}
  \item Assumptions on hydrodynamics:
    \begin{enumerate}
      \item Ideal fluid with TM EoS (\citealt{Taub}; \citealt{TM_EOS}).
    \end{enumerate}
%  \item Assumptions on cosmic-ray:
%    \begin{enumerate}
%      \item We treat CRs as a second fluid and solve directly for the evolution of CR pressure $p_{\text{cr}}$\
%            as a function of $\mathbf{r}$ and $t$.
%      \item We did not model the CR energy spectrum.
%      \item We neglected the cooling and heating processes of CRs, such as energy losses due to synchrotron\
%            and inverse Compton emission, and reacceleration in shocks.
%      \item We have assumed cosmic-ray is passive. (i.e. $p_{\text{cr}}\ll p_{\text{gas}}$)
%      \item \textcolor{red}{We have ignored the $\mathbf{B}$ field within\
%            the simulation box as the field inside\
%            the bubbles should be weak due to adiabatic expansion, and thus the magnetic fields has\
%            little effect on the overall dynamics.\
%            We also assumed the magnetic field is highly entangled on small-scale (how small?), resulting in\
%            the negligible cosmic-ray diffusion.}
%    \end{enumerate}
  \item Assumptions on gravity:
    \begin{enumerate}
      \item We use the potential of isothermal slab, symmetric about $z=0$ and\
            supported by gas pressure and by under its own self-gravity,\
            to mimic the realistic potential of Galactic bulge.
      \item We still use Newtonian gravity, as the relativistic fluid ejected by the jet source\
            is quickly slowed down by heavy disk in a short time, and the relativistic fluid\
            accounts for a little minority of total mass inside the simulation box.
      \item The ISM disk and atmosphere are subjected to the fixed external potential\
            due to a disk bulge and dark matter halo.
      \item In addition to the gravitational interaction, we ignore other interactions between stars and gases.
      \item We also ignore the self-gravity of the ISM disk and of the atmosphere.
      \item We ignore the centrifugal force of Milky Way rotation acting on the bubbles.
      \item We use the potential of isothermal slab to mimic the gravitational potential due to the stellar bulge.
      \item The interface between cold ISM disk and atmosphere is parallel to the Galactic plane and is pressure balanced.
    \end{enumerate}
\end{enumerate}
We simulate 3D special relativistic hydrodynamics with passive CR injections from the GC using\
the special hydrodynamics GPU code \textsc{gamer-sr} \citep{tseng2021}.

@@

\end{enumerate}
\subsection{Jet injection}
 \begin{enumerate}
    \item Why the overall structure is insensitive to jet direction?
      \begin{enumerate}
        \item Jet source size.
        \item Cold disk.
        \item Large pressure gradient along z-dircetion.
      \end{enumerate}
 \end{enumerate}

\subsection{X-ray and Gamma-ray emission}

  \begin{enumerate}
    \item X-ray:
       \begin{enumerate}
         \item Thermal bremsstrahlung: The X-ray emissivity in an energy range 1.4–1.6 keV is calculated\
               using the MEKAL model (\citealt{Xray-1}; \citealt{Xray-2}; \citealt{Xray-3})\
               implemented in the utility XSPEC\citep{XSPEC}, assuming solar metallicity.
       \end{enumerate}
    \item Gamma-ray:
       \begin{enumerate}
         \item Leptonic process:\\
          The gamma-ray emission is produced by inverse Comptom scattering of the ISRF by CRe.
         \item Hadronic process:\\
          In the hadronic model, CRp undergo hadronic collisions with thermal gas protons\
          and produce $\gamma$-ray via pion decay. The volume emissivity of the emission can be written as
          \begin{equation}
             \epsilon \propto U_{\text{CRp}}n_{p}\sigma_{p}\kappa_{pp}.
          \end{equation}
       \end{enumerate}
  \end{enumerate}

@@

Inverse Compton scattering
\begin{equation}
  \frac{dE}{dtd\epsilon_{1}dV} = \frac{3}{4}\sigma_{T}c\textit{C}\int^{\epsilon_{\text{max}}}_{\epsilon_{\text{min}}}
               \epsilon_{1}\frac{n(\epsilon)}{\epsilon}d\epsilon\int^{\gamma_{\text{max}}}_{\gamma_{\text{min}}}
               \gamma^{-(p+2)}f(q, \Gamma)d\gamma.
\end{equation}

\begin{equation}
f(q, \Gamma) = 2q\ln q+(1+2q)(1-q)+0.5(1-q)\frac{\left(\Gamma q\right)^2}{1+\Gamma q},
\end{equation}

\begin{equation}
\Gamma=\frac{4\epsilon \gamma_{\text{e}}}{m_{\text{e}}c^2}
\end{equation}

\begin{equation}
q=\frac{\epsilon_{1}/\gamma_{\text{e}}m_{\text{e}}c^{2}}{\Gamma\left(1-\epsilon_{1}/\gamma_{\text{e}}m_{\text{e}}c^{2}\right)}.
\end{equation}

Hadronic process
\begin{equation}
  \frac{dE}{dtd\epsilon_{1}dV} = cn_{H}pC\left(\frac{\epsilon_{1}}{m_{\text{p}}c^2}\right)^{-p}\int_{0}^{1}\sigma(\epsilon_{\text{p}}) F(x,\epsilon_{\text{p}}) x^{p}dx.
\end{equation}

\begin{equation}
\sigma(\epsilon_{\text{p}})=34.3+1.88L+0.25L^{2} \text{ mb}.
\end{equation}

\begin{equation}
F(x,\epsilon_{\text{p}})=B\frac{d}{dx}\left[\ln(x)\left(\frac{1-x^{\beta}}{1+kx^{\beta}\left(1-x^{\beta}\right)}\right)^4\right],
\end{equation}

where $B=1.30+0.14L+0.011L^2$, $\beta=\left(1.79+0.11L+0.008L^{2}\right)^{-1}$,\
$\left(0.801+0.049L+0.014L^{2}\right)^{-1}$, $L=\ln(\epsilon_{\text{p}}/1 \text{ TeV})$.

\section{to-do-list}
\begin{enumerate}
  \item Predict cold inner bubbles, emitting OIII or OII spectrum.
\end{enumerate}
